---
title: "Data overview"
format: html
execute: 
    echo: false
    include: false
    message: false
    warning: false
---

```{r}
library(DT)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(patchwork)
library(rstatix)
library(GGally)
library(ggcorrplot)
library(fixest)
library(here)
library(psych)
```

```{r}
# Data used in most plots (has life_expectancy, fertility_rate, etc.)
join_labeled <- readRDS(here("data", "processed", "join_labeled.rds"))

# Data used later in the FE models
join_clean   <- readRDS(here("data", "final", "join_clean.rds"))

# Data used in the Appendix.
# Has descriptive uses mostly in this notebook
wealth <- read.csv(here("data", "raw", "country_wealth_labels_meta.csv"))
# Rename columns for aethestics
colnames(wealth) <- append(c("labels"), sapply(seq(1987, 2024), FUN = as.character))
```


The main information available is composed yearly records of multiple indicators across multiple countries. In total, information on 19 indicators over 210 different countries from 2000 up to 2014 was recolted. This information can be downloaded publicly from the dedicated website [World Bank Open Data](https://data.worldbank.org/indicator) or through the R -`WDI`- package. Indicators include markers of economical and financial factors such as GDP, GPD growth, GNI, total governement reserved in USD or socio-economical factors such as population, enrollement in primary, secondary and tertiary schools or fertility rate. See [Appendix A1](#a1.-indicators) for an overview and description of every indicators.

As differentiation between 210 different countries can prove to be difficult. Fortunately, the World Bank Group provides a robust and systematic methodology for classifying countries by income level[^atlas-blog]. This classification is based on the Gross National Income in US dollar for each country. The classification differ every year. [Appendix A2](#a2.-wealth-classes-2000-2024) displays the full extrend of the income classes requirement for every year.

[^atlas-blog]: [A World Bank Blog article about the methodology](https://blogs.worldbank.org/en/opendata/world-bank-country-classifications-by-income-level-for-2024-2025)

```{r}
#| label: tbl-oveview-wealthclass
#| tbl-cap: Overview of the recent income classes
#| include: true
wealth %>% 
    select("labels", "2020":"2024") %>% 
    DT::datatable(
        options = list(
            searching = FALSE,
            paging = FALSE),
        class = "compact"
    )
```

The datasets counts 4 main identification variables: 
* Country
* Iso2c
* Iso3c
* Year

ISO code aren't used and will be left out for most of the report.

---

Find below short but usefull descriptive statistics of our key-indicators.

'N' symbolise the amount of 'usable' observation inside the variable. 
Unsable observation are missing values and the maximum amount of observations is **5250**. 

```{r}
#| label: tbl-overview-socio-economic
#| tbl-cap: Overview of socio-economic indicators
#| include: true
socioeco_vars <- c("unemployment_pct", "population_total", "urban_pop_pct", "life_expectancy", "net_enrol_primary", "net_enrol_secondary", "enrol_tertiary_gross", "fertility_rate", "pop_growth_pct")
join_labeled %>% 
    ## Select socio-economic variables only
    select(all_of(socioeco_vars)) %>% 
    ## psych::describe()
    describe(.) %>%
    ## Filter out unwanted descriptive statistics
    select(n, mean, median, sd, min, max) %>% 
    ## Rounding for visibility
    mutate_all(.funs = ~ round(., digits = 3)) %>% 
    ## DT::datatable for aethestics and customization
    DT::datatable(
        options = list(
            searching = FALSE,
            paging = FALSE),
        colnames = c("N", "Mean", "Standard deviation", "Median", "Minimum", "Maximum"),
        rownames = c("Unemployement (%)", "Population total", "Urban Population (%)", "Life expectancy", "Primary education enrol.", "Secondary education enrol.", "Tertiard enducation enrol. gross", "Fertility rate", "Population growth (%)"),
        class = "compact")
```

```{r}
#| label: tbl-overview-economic
#| tbl-cap: Overview of economic indicators
#| include: true
id_vars = c("country", "iso2c", "iso3c", "year")
join_labeled %>% 
    ## Select economic variables only
    select(!c(id_vars, socioeco_vars)) %>% 
    ## psych::describe()
    describe(.) %>%
    ## Filter out unwanted descriptive statistics
    select(n, mean, median, sd, min, max) %>% 
    ## Rounding for visibility
    mutate_all(.funs = ~ round(., digits = 3)) %>% 
    ## DT::datatable for aethestics and customization
    DT::datatable(
        options = list(
            searching = FALSE,
            paging = FALSE),
        colnames = c("N", "Mean", "Standard deviation", "Median", "Minimum", "Maximum"),
        rownames = c("Inflation (%)", "Dept (% of GDP)", "GDP growth (%)", "GDP ($)", "Exports (% of GDP)", "Imports (% of GDP)", "Broad Money (%)", "Total reserve ($)", "GNI ($)", "Atlas Conversion Factor", "Income class"),
        class = "compact")
```

For more insights into the joining process, refer to [data_joining](data_joining.qmd).

## Missing Values

Some indicators contain a lot of missing values. Handling missing values should be done with care, as it influences the quality of the overall dataset and thus the prediction. Indeed, in our case removing rows containing missing values could bias the analysis, as the proportion of missing values in country with lower income is different than in country with higher income. Additionnaly, removing entire column could prove detrimental to the overall analysis as we could inadvertently remove important informations. Thus, the removal strategy will depend in the analysis performed.

In the graph below, you can observe the map representation of those missing values, across all countries and summed up over the years.

Since our time granularity is year, this means that countries in red provide very few information at all. Additionnaly, even though european countries seems to contain less missing values than others countries, they are not exempt of them.
Some indicators seems less present in poorer countries. However, other factors could influence the presence of missing values other the the poverty of the country. For example, strict and opaque political regime, such as China, are less likely to divulge truthful information, if any at all.

```{r}
#| include: true
#| label: fig-worldplot
#| fig-cap: Worldplot of missing value for GDP

## Number of total missing value per country per year
na_count <- join_labeled %>%
    ## Transform in longer fromat for easier NaN's count computation
    pivot_longer(
        cols = inflation_pct:pop_growth_pct,
        names_to = "indicator",
        values_to = "indic_values") %>% 
    group_by(country, indicator) %>% 
    ## detect NA values per country, year and indicator
    mutate(
        is_na = is.na(indic_values),
        na_count = sum(is_na),
        country = as.factor(country),
        country = recode(country, "United States" = "USA", "Russian Federation" = "Russia")) ## Renaming for consistency


## Get world's countries coordinates
map <- map_data("world") %>% 
    ## Join countries coordinates with NaN's summary
    ## Inner-join to rule out discrepancy between 'world' and 'is_na'
    inner_join(
        .,
        na_count %>% filter(indicator == "debt_gdp_pct"),
        by=c("region"="country"),
        relationship = "many-to-many")

## Create a map of the world, filled by amount of NaN's per countries
plain <- theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    panel.background = element_rect(fill = "white"))

## Create a map of the world, filled by amount of NaN's per countries
worldplot <- map %>% 
    ggplot(
        data = .,
        mapping = aes(x = long, y = lat)) +
    coord_fixed(1.3) +
    geom_polygon(
        mapping = aes(fill = na_count, group = group)) +
    scale_fill_distiller(palette = "RdBu", name = "NaN's Amount") +
    ggtitle("Total missing value in Dept (% of GDP)") +
    labs(caption = "Remark, divergence between country name might create blank values") +
    plain

ggplotly(worldplot)
```