# Data cleaning

The data available comprised of multiple variables having a lot of missing values. 
In the exploration phase, we could clean missing values on the go due to the limited scope of the variable involved. 
However, modelling inflation using many estimators requires a cleaning steps before hand.

Removing rows or columns should be done with care as it could create biases.

During the overview of the data in the report, we identified variables comprising a lot of different missing values.
Additionnaly, we also previously identified that most of those missing values figures in poorer countries.
```{r}
#| include: true
join_labeled %>% 
    summarise(across(everything(), ~ sum(is.na(.)))) %>% 
    select(-country, -iso2c, -iso3c, -year) %>% 
    pivot_longer(1:20, names_to="indicators", values_to="nan") %>% 
    mutate(nan_pct = round(nan / dim(join_labeled)[1] * 100, digits = 0)) %>% 
    knitr::kable(.)
```

We want to indentify predictors for inflation rate across the rate without limiting the sample to rich country only. In this sense, we need to keep the dataset as wide and long as possible, at least for a first modelization. Thus, we'll drop out variables having a majority of missing values, and simply remove the rest row-wise. Although we could avoid removing some variables by conditioning on a specific population, said richer countries for example, we chose to look at inflation rate at its globality. Keeping those variable would introduce a clear sampling bias, hence why we remove them.

`dept_gdp_pct` is on the limit, however its an important economical factor describing the general healt of an economy. Consequently, we'll keep it.

Other indicators present strong collinearity with each other, hence, we'll have to drop them to present the integrity of the prediction.
More specifically, `imports_gdp_pct` will removed. Choice for keeping exports information rather than imports is arbitrary.
`gni_usd` will also be remove since it transmits the same information as `gdp_usd` and it is an extension of the later since it is computed based on it.

List of variable that will be removed is: 
* `life_expectancy` with 74% missing values
* `atlas_conversion_factor` with 55% missing values
* `fertility_rate` with 65% missing values
* `population_total` with 43% missing values
* `imports_gdp_pct` for collinearity
* `gni_usd` for collinearity

```{r}
join_labeled_clean <- join_labeled %>% 
    select(
        -life_expectancy,
        -altas_conversation_factor,
        -fertility_rate,
        -population_total,
        -imports_gdp_pct,
        -gni_usd) %>% 
    drop_na()
```

Because of the amount predictors we would like to inspect, the cleaning reduced drastically the size of the dataset from `{r} dim(join_labeled)[1]` to `{r} dim(join_labeled_cleaned)[1]`. Although we managed to preserve a relatively unbiased population, the poorest country are almost absent of the dataset. This an important point to keep in mind.

```{r}
print("List of country in the 'clean' dataset.")
unique(join_labeled_clean$country)

print("Proportion of wealth classes in the dataset")
join_labeled_clean %>% 
    group_by(wealth_class) %>% 
    summarize(value_counts = n())

print("'Poor' country: ")
join_labeled_clean %>% 
    filter(wealth_class == "L") %>% 
    select(country) %>% 
    unique(.)
```