---
title: "Inflation Developement"
subtitle: "An exploration of patterns and predictors"
author:
    - name: Girardin Robin
    - name: Geogantopoulos Nikolaos
institute: "G & G Data Science Cabinet"
execute: 
    echo: false
    include: false
    message: false
    warning: false
    cache: true
format: 
    html: 
        toc: true
        toc-float: true
---

Inflation influences consumer purchasing power, economic growth, buisinesses developpement and competiteveness, central banks interests rates. As such, understand such an important factor of nowadays economy is crucial. Even more so considerinng the actual political and economical turbulence.

We were mandated by [World Bank Group](#https://www.worldbank.org/ext/en/home), Departement of Developement Economics, to explore data at their disposal and determine inflation developement patterns and predictors. More specifically, this report targets the following questions: 

- How is inflation distributed?
- How does inflation develops from 2000 until todays?
- How does crisis influence inflation markers?
- Which indicators could be used to predict inflation rates?

```{r}
library(DT)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(patchwork)
library(rstatix)
```

```{r}
join_labeled <- readRDS("/Users/robingirardin/school/inflation-developement/data/processed/join_labeled.rds")
```


# Overview of the data

The main information available is composed yearly records of multiple indicators across multiple countries. In total, information on 19 indicators over 210 different countries from 2000 up to 2014 was recolted. This information can be downloaded publicly from the dedicated website [World Bank Open Data](https://data.worldbank.org/indicator) or through the R -`WDI`- package. Indicators include markers of economical and financial factors such as GDP, GPD growth, GNI, total governement reserved in USD or socio-economical factors such as population, enrollement in primary, secondary and tertiary schools or fertility rate. See [Appendix A1](#a1.-indicators) for an overview and description of every indicators.

As differentiation between 210 different countries can prove to be difficult. Fortunately, the World Bank Group provides a robust and systematic methodology for classifying countries by income level[^atlas-blog]. This classification is based on the Gross National Income in US dollar for each country. The classification differ every year. [Appendix A2](#a2.-wealth-classes-(2000-2024)) displays the full extrend of the income classes requirement for every year.

[^atlas-blog]: [A World Bank Blog article about the methodology](https://blogs.worldbank.org/en/opendata/world-bank-country-classifications-by-income-level-for-2024-2025)


| Label | 2022 | 2023 | 2024 |
|-------|------|------|------|
| High (H) | <= ,.135 | <= 1,145 | <= 1,135 |
| Upper Middle (UM) | 1,136 - 4,465| 1,146 - 4,515 | 1,136 - 4,495 |
| Lower Middle (LM) | 4,466 - 13,845 | 4,516 - 14,005 | 4,496 - 13,935 |
| Low (L) | > 13,845 | > 14,005 | > 13,935 |

: Wealth classes {#tbl-wealth-meta}

::: {#tbl-panel layout-ncol=2}
| Variable | Type | NaN | Unique |
|----------|------|-----|--------|
| country  | Categorical | 0 | x |
| iso2c    | Categorical | 0 | x |
| iso3c    | Categorical | 0 | x |
| wealth_class | Categorical | 0 | 4 |

: Identifier {#tbl-first}

| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: Economical Indicator {#tbl-second}

| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: Socio-economial Indicator {#tbl-second}

Data overview
:::

For more insights into the joining process, refer to [data_joining](data_joining.qmd).

## Missing Values

As @tbl-panel shows, some indicators contain a lot of missing values. Handling missing values should be done with care, as it influences the quality of the overall dataset and thus the prediction. Indeed, in our case removing rows containing missing values could bias the analysis, as the proportion of missing values in country with lower income is different than in country with higher income. Additionnaly, removing entire column could prove detrimental to the overall analysis as we could inadvertently remove important informations. Thus, the removal strategy will depend in the analysis performed.

In the graph below, you can observe the map representation of those missing values, across all countries and summed up over the years.

Since our time granularity is year, this means that countries in red provide very few information at all. Additionnaly, even though european countries seems to contain less missing values than others countries, they are not exempt of them.
Some indicators seems less present in poorer countries. However, other factors could influence the presence of missing values other the the poverty of the country. For example, strict and opaque political regime, such as China, are less likely to divulge truthful information, if any at all.

```{r}
#| include: true
## Number of total missing value per country per year
na_count <- join_labeled %>%
    ## Transform in longer fromat for easier NaN's count computation
    pivot_longer(
        cols = inflation_pct:pop_growth_pct,
        names_to = "indicator",
        values_to = "indic_values") %>% 
    group_by(country, indicator) %>% 
    ## detect NA values per country, year and indicator
    mutate(
        is_na = is.na(indic_values),
        na_count = sum(is_na))

## Get world's countries coordinates
map <- map_data("world") %>% 
    ## Join countries coordinates with NaN's summary
    ## Inner-join to rule out discrepancy between 'world' and 'is_na'
    inner_join(
        .,
        na_count %>% filter(indicator == "debt_gdp_pct"),
        by=c("region"="country"),
        relationship = "many-to-many")

## Create a map of the world, filled by amount of NaN's per countries
plain <- theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    panel.background = element_rect(fill = "white"))

    ## Create a map of the world, filled by amount of NaN's per countries
worldplot <- map %>% 
    ggplot(
        data = .,
        mapping = aes(x = long, y = lat)) +
    coord_fixed(1.3) +
    geom_polygon(
        mapping = aes(fill = na_count, group = group)) +
    scale_fill_distiller(palette = "RdBu", name = "NaN's Amount") +
    ggtitle("Total missing value in Dept (% of GDP)") +
    plain

ggplotly(worldplot)
```

# Exploration

Our main point of focus is the inflation rate. Thus, let's plot this for all available countries first. Luckily, this variable hardly contain any missing values at all (`{r} round(sum(is.na(join_labeled$inflation_pct)) / length(join_labeled$inflation_pct), digits = 3)` - `{r} sum(is.na(join_labeled$inflation_pct))`). Hence, we can just leave them aside for now.

Looking at inflation broadly, we observe a high amount of extreme values and a huge concentration around 0. Ultimately, this also means that we don't see much of information here. Hence, transforming the data into some more normally distributed, and thus reducing the influence of extreme values might be usefull. The plot below illustrate that a logarithmic transformation works particularly well here.

```{r}
#| include: true
p1 <- ggplot(
    data = join_labeled,
    mapping = aes(x = year, y = inflation_pct)) +
        geom_point(color = "#2b8cbe", alpha = 0.3) +
        geom_smooth(color = "black", method="lm") +
        geom_smooth(color = "blue", method = "gam") +
        labs(x = "Year", y = "Inflation (%)") +
        theme_minimal()

p2 <- ggplot(
    data = join_labeled,
    mapping = aes(x = year, y = log(inflation_pct))) +
        geom_point(color = "#2b8cbe", alpha = 0.3) +
        geom_smooth(color = "black", method="lm") +
        geom_smooth(color = "blue", method = "gam") +
        labs(x = "Year", y = "Ln(Inflation) (%)") +
        theme_minimal()

p3 <- ggplot(
    data = join_labeled,
    mapping = aes(x = inflation_pct)) +
        geom_density(fill = "#2b8cbe", color = "#2b8cbe", alpha = 0.3) +
        labs(x = "Inflation (%)", y = "Density)") +
        theme_minimal()

p4 <- ggplot(
    data = join_labeled,
    mapping = aes(x = log(inflation_pct))) +
        geom_density(fill = "#2b8cbe", color = "#2b8cbe", alpha = 0.3) +
        labs(x = "Ln(Inflation) (%)", y = "Density") +
        theme_minimal()

(p1 + p2) / (p3 + p4) + plot_annotation(title = "Inflation and log-transformed inflation distribution")
```

Here, we can observe that, although inflation rate was relatively low across the past 25 years, some countries still experienced and are still experiencing comparatively high inflation (\~20 %) compared to other countries. Additionnaly the richest the country, the less inflation it has.

Concerning trend, poorest countries (L & LM) demonstrate a clear overall pattern across the past 25 years. Their inflation rate diminished up around 2010 before increasing again. Upper-middle countries maintained a slow but steady inflation increase, while wealthier countries show a more distributed pattern. They displayed a cycle pattern of global inflation increase at major crisis (2008 - subprime crisis and 2020 - COVID crisis), before entering a deflation phase. Interestingly, only wealthy countries seemed affect by the 2008 and 2020 crisis, as if the term "Global" in Global Financial crisis don't accounts for less developed countries. This phenomenon could be due to the effect of outliers biasing the trend. Indeed, removing the impact of outliers by transforming the data, we can observe a more realistic scenario with a more gradual pattern distinction between income class. Indeed, although each classes harbors those cyclical pattern around major crisis, the extend to which their influence varies across income class. The richer the country, the more influenced they are. This could be partly explained by the highly globalized econonmy of richer country during the past years. Indeed, the richer the country, the more heavily they rely on the neighbors, the bigger the influence of global crisis are, compared to poorer country which are outside of such a globalized world.

```{r}
#| include: true
dataplot_wealthclass_inflation <- join_labeled %>% 
    # Select only the necessary columns before dropping NaN's
    select(year, country, wealth_class, inflation_pct) %>% 
    # Avoid 'NA' income class group.
    drop_na()

deviation_summary <- dataplot_wealthclass_inflation %>% 
    group_by(wealth_class) %>% 
    summarize(
        mean = mean(inflation_pct),
        standard_deviation = sd(inflation_pct),
        quartile1 = quantile(inflation_pct, 0.25),
        quartile3 = quantile(inflation_pct, 0.75)
    )
knitr::kable(deviation_summary, caption = "Important spread statistics of inflation")
```

On top of the fact that the richer countries, the lower rate of inflation on average, inflation rate seems to vary greatly between the difference income classes. Indeed, the richer the country, the less its inflation rate varies.  Smaller rates of inflation are linked with a growing and heathly economy. Hence, the healthier the economy, the stabler it gets. 
This stability difference could be biased by the amount of extreme values of inflation rate in poorer country. Indeed, the transformed graph shows much less variation between each income class. However, a more precise look at the distribution across income classes for the logarithmically transformed inflation would strill sreveal differences in shape and standard deviation.

```{r}
#| include: true

p5 <- ggplot(
    data = dataplot_wealthclass_inflation,
    mapping = aes(x = year, y = inflation_pct, color = wealth_class)) +
        geom_smooth() +
        labs(y = "Inflation (%)", subtitle = "Normal", color = "Income class") +
        scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
        theme_minimal()

p6 <- ggplot(
    data = dataplot_wealthclass_inflation,
    mapping = aes(x = year, y = log(inflation_pct), color = wealth_class)) +
        geom_smooth() +
        labs(y = "Ln(Inflation) (%)", subtitle = "Logarithmically transformed", color = "Income class") +
        scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
        theme_minimal()

p5 / p6 + plot_annotation("Inflation developement across income class")
```

```{r}
#| include: true
p6 <- ggplot(
    data = dataplot_wealthclass_inflation,
    mapping = aes(x = log(inflation_pct), fill = wealth_class)) +
        geom_density(alpha = 0.3) +
        labs(x = "Ln(Inflation) (%)", y = "Density", fill = "Income Class") +
        theme_minimal()

ggplotly(p6)
```

## Statistical Analysis

### Do richer country have lower inflation rate?

Say that the inflammation rate of each country is a random variable $X$ of 25 observation recolted over the years such that:

$$
X = \{x_1, x_2, x_3, \dots, x_{25}\}
$$

$X$ is auto-correlated since $x_i$ depends on previous values. Indeed, if $x_{22}$ is high, there is a higher chance that $x_{23}$ will high as well. As such, $X$ isn't independent.

```{r}
#| include: true
ch_acf <- acf(
    join_labeled %>% filter(country == "Switzerland") %>% select(inflation_pct),
    plot = FALSE)

print("Auto-correlation of Switzerland's inflation a different lag year:")
print(ch_acf)
```

Additionnaly, it is hard to assume that the inflation rate of each country is independant on another. A particular example is demonstrated below.

```{r}
#| include: true
ch_fr_cor <- cor(
    join_labeled %>% filter(country == "Switzerland") %>% select(inflation_pct),
    join_labeled %>% filter(country == "France") %>% select(inflation_pct),
)

print(paste0("Switzerland - France correlation: ", round(ch_fr_cor[1, 1], digits = 3)))
```

More specifically, the mean correlation between the inflation rate of all rich country is 0.44. It is enough for us to be prudent and assume dependance on this side.

```{r}
#| include: true

wide_country <- join_labeled %>% 
    filter(wealth_class == "H") %>% 
    select(inflation_pct, year, country) %>%
    drop_na() %>%
    group_by(year) %>% 
    mutate(mean = mean(inflation_pct)) %>% 
    pivot_wider(id_cols = year, names_from = country, values_from = mean, values_fill = 0) %>% 
    ungroup() %>% 
    select(-year)

cor_matrix <- expand.grid(colnames(wide_country), colnames(wide_country))

correlation = c()
for (i in 1:4761){
    row = cor_matrix[i, ]
    x_country = row[["Var1"]]
    y_country = row[["Var2"]]
    x = join_labeled %>% filter(country == x_country) %>% select(inflation_pct)
    y = join_labeled %>% filter(country == y_country) %>% select(inflation_pct)
    p = cor(x[, 1], y[, 1], use="complete")
    correlation = append(correlation, p)
}

cor_matrix = cbind(cor_matrix, correlation)

mean_cor <- mean(cor_matrix[cor_matrix$correlation < 1, ]$correlation)

print(paste0("Mean pearson correlation between countries: ", round(mean_cor, digits=2)))

```

Nonetheless, we try our best to meet this condition by randomly taking a sample of observation in each group. This way, we reduce the possibility of cross and within-country influence on the inflation rate. This case is a classic case of an unpaired test, testing difference in means. A common test for this scenario would be an analysis of variance (ANOVA) or non-parametric tests.

In our example, as demonstrated by the plots below the data isn't normally distributed. In fact, the distribution looks very similary to the gamma distribution. Mapping the inflation rate to all countries with a gamma distribution $\Gamma(3, 0.5)$, we can see the heavy similarity between the two. Since the observation of each wealth class seems to come from a similar gamma distribution, the assumption of identically distributed observations is statisfied. Hence, a sensible choice would be to go for a non-parametric tests, less sensible to outlier and extreme values. More specifically, a Kruskal-Wallis test since we 4 independent group: "H", "UM", "LM" and "L".

```{r}
#| include: true

ww_density <- ggplot(
    data = dataplot_wealthclass_inflation,
    mapping = aes(x = inflation_pct)) +
    geom_density(color = "#addd8e", size = 0.75) +
    labs(x = "Inflation (%)", y = "Density") +
    ggtitle("'World-wide Inflation Distribution") +
    theme_minimal()

cls_density <- ggplot(
    data = dataplot_wealthclass_inflation,
    mapping = aes(x = inflation_pct, color = wealth_class)) +
    geom_density(size = 0.75) +
    xlim(0, 400) +
    labs(x = "Inflation (%)", y = "Density") +
    ggtitle("Inflation Distribution by Wealth Class") +
    theme_minimal() + 
    facet_wrap(~wealth_class)

print(ww_density)
print(cls_density)
```

```{r}
#| include: true

gamma <- rgamma(length(dataplot_wealthclass_inflation$inflation_pct), 3, rate=0.5)
ww_density_gamma <- ggplot(
    data = dataplot_wealthclass_inflation) +
    geom_density(
        mapping = aes(x = inflation_pct, color = "Inflation"), 
        size = 0.75) +
    geom_density(
        mapping = aes(x = gamma, color = "Gamma"), 
        size = 0.75) +
    ggtitle("Goodness of fit") + 
    scale_color_manual(values = c("Inflation" = "#addd8e", "Gamma" = "#2c7fb8")) +
    labs(x = "Inflation (%)", y = "Density", color = "Distribution") +
    theme_minimal()
print(ww_density_gamma)
```

As per the Kruskal-Wallis Test, we observe strong evidence against the null hypothesis that the median inflation difference between income classes is not 0. More specifically, we observe strong evidence against the null hypothesis that the median inflation difference of each income class is not 0. In other words, each wealth class are significantly different from each other. Since the sample size is big, the effect size is a biased statistic, hence why we compute the effect size alongside it. Said effect size is considered large and thus support the previous significant claims.
More importantly, the same test performed on the logarithmic transformed inflation, less biased by outlier deliver similar results.

**Conclusion:** The richer the country, the lower is the inflation rate. **However**, it is important to note that such insights need to be taken with a grain of salt, as observation aren't totally independant of each other.

```{r}
#| include: true
random_sample <- join_labeled %>% 
    select(wealth_class, country, inflation_pct) %>% 
    drop_na() %>% 
    group_by(wealth_class) %>% 
    slice_sample(prop=0.25) %>% 
    ungroup()

kw_model_normal <- kruskal.test(inflation_pct ~ wealth_class, data = random_sample)
kw_p_normal <- kw_model_normal$p.value

kw_eta_normal <- random_sample %>% 
    kruskal_effsize(inflation_pct ~ wealth_class)

dunn_post_hoc <- dunn_test(random_sample, inflation_pct ~ wealth_class)

print(paste0("Kruskal-Wallis Test p-value: ", kw_p_normal))
print(
    paste0(
        "Kruskal-Wallis Test effect size: ",
        round(kw_eta_normal$effsize, digits = 3),
        " with a ",
        kw_model_normal$magnitude,
        " magnitude"))

knitr::kable(dunn_post_hoc %>% select(.y., group1, group2, p.adj, p.adj.signif))
```

```{r}
kw_model_log <- kruskal.test(log(inflation_pct) ~ wealth_class, data = random_sample)
kw_p_log <- kw_model_log$p.value

kw_eta_normal <- random_sample %>% 
    kruskal_effsize(inflation_pct ~ wealth_class)

print(paste0("Kruskal-Wallis Test p-value: ", kw_p_log))
print(
    paste0(
        "Kruskal-Wallis Test effect size: ",
        round(kw_eta_normal$effsize, digits = 3),
        " with a ",
        kw_eta_normal$magnitude,
        " magnitude"))
```

### Do richer country have a more stable inflation rate compared to poorer countries?

In [Inflation trend per wealth class](#inflation-trend-per-wealth-class) we've observed graphically the variation difference in inflation between wealth classes (see @intra-class-inflation-var).

In order to statistically confirm whether this assumption is right or wrong, we'll measure the standard deviation difference between each class.

```{r}
#| include: true

join_labeled %>% 
    select(wealth_class, inflation_pct, year) %>% 
    drop_na() %>% 
    group_by(wealth_class) %>% 
    summarise(
        sd = sd(inflation_pct),
        log_sd = sd(log(inflation_pct), na.rm = TRUE)) %>% 
    knitr::kable()
```

Similar to the [Previous Section](#do-richer-country-have-lower-inflation-rate), we don't consider the inflation rate of a country to be independent, as proven by the previously demonstrated by the slight auto-correlation. Hence, the same could be said by their standard deviation. Additionnaly, it is also unlikely for the inflation rate between countries to be independent. Consequently, we'll select randomly a sample from each sample, across all countries and years measured. This will ensure minimum influence of countries between each other and reduce the auto-correlation between observations of the same countries. The latter is further supported by the fact that said auto-correlation becomes negligible after 2-3 years.

 Since the data isn't normally distributed, we'll use the non-parametric Kruskal-Wallis Test.

As per the Kruskal-Wallis test, we observe strong evidence against the null hypothesis that the standard deviation inflation difference between income classes is not 0. More specifically, we observe strong evidence against the null hypothesis that the standard deviation inflation difference of each income class is not 0, except between low and middle low income country. In other words, the wealth class has a significant effect on the inflation standard deviation, with a large effect size. Each wealth class standard deviation is significantly different from each other.

**Conclusion:** The richer the country, the stabler is the inflation rate across country. For the richer country, which experience high perturbation during the 2008 and 2020 crisis, this means that, although the inflation rate changed, everything country moved in the same direction. More importantly, a similar analysis on logarathmic transformed inflation rate, reducing the impact of outlier, deliver similar result. **However**, it is important to note that such insights need to be taken with a grain of salt, as observation are totally independant of each other.

This could be the proof, or the consequence, or a highly globalized economy where each country is dependent on each other. As per the recent global economical turmoil, this can signify global drawbacks.

```{r}
#| include: true

random_sample <- join_labeled %>% 
    select(wealth_class, country, inflation_pct) %>% 
    drop_na() %>% 
    group_by(wealth_class) %>% 
    slice_sample(prop=0.25) %>% 
    mutate(
        sd_infl_country = inflation_pct,
        sd_infl_country_log = sd(inflation_pct), na.rm = TRUE) %>% 
    ungroup() 

kw_model_normal<- kruskal.test(sd_infl_country ~ wealth_class, data = random_sample)
kw_p_normal <- kw_model_normal$p.value

kw_eta_normal <- random_sample %>% 
    kruskal_effsize(sd_infl_country ~ wealth_class)

dunn_post_hoc <- dunn_test(random_sample, sd_infl_country ~ wealth_class)

print(paste0("Kruskal-Wallis Test p-value: ", kw_p_log))
print(
    paste0(
        "Kruskal-Wallis Test effect size: ",
        round(kw_eta_normal$effsize, digits = 3),
        " with a ",
        kw_eta_normal$magnitude,
        " magnitude"))

knitr::kable(dunn_post_hoc %>% select(.y., group1, group2, p.adj, p.adj.signif))
```


```{r}
kw_model_log <- kruskal.test(sd_infl_country_log ~ wealth_class, data = random_sample)

kw_eta_log <- random_sample %>% 
    kruskal_effsize(sd_infl_country_log ~ wealth_class)

kw_p_log <- kw_model_log$p.value
dunn_post_hoc <- dunn_test(random_sample, sd_infl_country_log ~ wealth_class)

print(paste0("Kruskal-Wallis Test p-value: ", kw_p_log))
print(
    paste0(
        "Kruskal-Wallis Test effect size: ",
        round(kw_eta_log$effsize, digits = 3),
        " with a ",
        kw_eta_log$magnitude,
        " magnitude"))
```


# Disclamer on GenAI


# Source

Indicator dataset:

1. [World Bank Group](https://data.worldbank.org/indicator)
2. [WDI](https://cran.r-project.org/web/packages/WDI/WDI.pdf)
3. ['join.rds'](data/processed/join.rds)

Wealth classification dataset: 

1. [Wealth Classification](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)

# Appendix

## A1. Indicators

Find below a short description of each indicators composing the main dataset. The description has the following format

* `WDI` indicator code (human readable label): description

**Description:**

* FP.CPI.TOTL.ZG (inflation_pct): Inflation, consumer prices (annual %) — measures the yearly percentage change in consumer prices as measured by the consumer price index. 

* GC.DOD.TOTL.GD.ZS (debt_gdp_pct): Central government debt, total (% of GDP) — the stock of direct government debt liabilities as a share of gross domestic product.

* NY.GDP.MKTP.KD.ZG (gdp_growth_pct): GDP growth (annual %) —  the total income earned through the production of goods and services in an economic territory during an accounting period. Measured in USD.  

* NY.GDP.MKTP.CD (gdp_usd): GDP (current US$) — the total monetary value of all goods and services produced in an economy, measured at current U.S. dollar exchange rates. 

* SL.UEM.TOTL.ZS (unemployment_pct): Unemployment, total (% of labor force) — the share of the labor force that is without work but available for and seeking employment.  

* NE.EXP.GNFS.ZS (exports_gdp_pct): Exports of goods and services (% of GDP) — exports of goods and non-factor services as a percentage of GDP.

* NE.IMP.GNFS.ZS (imports_gdp_pct): Imports of goods and services (% of GDP) — imports of goods and non-factor services as a percentage of GDP.

* SP.POP.TOTL (population_total): Population, total — the total population of a country.

* FM.LBL.BMNY.GD.ZS (broad_money_pct): Broad money (% of GDP) — the sum of all liquid financial instruments held by money-holding sectors that are widely accepted in an economy as a medium of exchange expressed as a share of GDP. 

* FI.RES.TOTL.CD (total_reserve_usd): Reserves and related items (total reserves in current US$) — the stock of a country’s foreign exchange and gold reserves. 

* NY.GNP.MKTP.PP.CD (gni_usd): GNI, PPP (current international $) — gross national income converted to international dollars using purchasing power parity (PPP) conversion factors. 

* PA.NUS.PPP (altas_conversion_factor): Atlas conversion factor — World Bank’s Atlas method to convert national currencies into U.S. dollars to smooth exchange rate fluctuations. 

* SP.URB.TOTL.IN.ZS (urban_pop_pct): Urban population (% of total) — the percentage of a country’s population living in urban areas.

* SP.DYN.LE00.IN (life_expectancy): Life expectancy at birth, total (years) — the average number of years a newborn is expected to live if current mortality rates continue to apply.

* SE.PRM.NENR (net_enrol_primary): School enrollment, primary (% net) — net enrollment rate in primary education: pupils of official primary school age enrolled as a proportion of the total population of that age.

* SE.SEC.NENR (net_enrol_secondary): School enrollment, secondary (% net) — net enrollment rate in secondary education. 

* SE.TER.ENRR (enrol_tertiary_gross): School enrollment, tertiary (% gross) — the gross enrollment ratio for tertiary education: total enrollment in tertiary education, regardless of age, expressed as a percentage of the official tertiary education age. 

* SP.DYN.TFRT.IN (fertility_rate): Fertility rate, total (births per woman) — the average number of children that would be born to a woman over her lifetime if she were to experience the age-specific fertility rates of a given period.  

* SP.POP.GROW (pop_growth_pct): Population growth (annual %) — the annual percentage growth rate of the total population, including the effects of natural increase and net migration.

## A2. Wealth classes (2000-2024)

Find below the classification rule for the income level classes from 2000 until 2024.

* L: low income class
* LM: lower middle income class
* UM: upper middle income class
* H: High income class

```{r}
#| include: true
wealth <- read.csv("/Users/robingirardin/school/inflation-developement/data/raw/country_wealth_labels_meta.csv")

# Remove the 'X' in the years column
colnames(wealth) <- append(c("labels"), sapply(seq(1987, 2024), FUN = as.character))

wealth %>% 
    select("labels", "2000":"2024") %>% 
    DT::datatable(
        .,
        options = list(
            dom = 't'
        ))
```
