---
title: "Prediction inflation rate"
format: html
execute: 
    echo: false
    include: false
    message: false
    warning: false
---

```{r}
library(DT)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(patchwork)
library(rstatix)
library(GGally)
library(ggcorrplot)
library(fixest)
library(here)
library(psych)
```

```{r}
# Data used in most plots (has life_expectancy, fertility_rate, etc.)
join_labeled <- readRDS(here("data", "processed", "join_labeled.rds"))

# Data used later in the FE models
join_clean <- readRDS(here("data", "final", "join_clean.rds"))

# Data used in the Appendix.
# Has descriptive uses mostly in this notebook
wealth <- read.csv(here("data", "raw", "country_wealth_labels_meta.csv"))
# Rename columns for aethestics
colnames(wealth) <- append(c("labels"), sapply(seq(1987, 2024), FUN = as.character))
```

Due to the fact that our data is panel data, we'll use fixed-effects models to account for time and country fixed effects. 

However, before modeling inflation rate based on multiple predictors, we'll need to clean the data of its missing values. 
In the exploration part, as we were exploring a handful of indicators at a time, we could do it case by case.
Now, however, as we're handling 19 differents indicators at the same time, we'll need a slightly more refined approach.
See [Data Cleaning](data_cleaning.qmd) for more information. The cleaning processed revealed a dataset with little representation of low income country. In order to avoid this, we could adopt a different missing value cleaning policy based on the result of the first model.

Additionnaly, in the [Exploration - Indicators](#indicators) section, we illustrated the benefits of using a log-transformation to certain indicators. We'll perform such an operation to economic indicators only, as they were the only revealing the most skewed and unfitting distribution for a linear model. Concerning socio-economical indicators, although still anormally distributed, it is still acceptable and we would want to artificially modify too many estimators as long as we don't need to. 
Another issue is that log-transformation doesn't work with negative values, which is in minority present in inflation rate. Luckily, filtering them out isn't conceptually wrong since we focus on 'inflation' and not 'deflation'.

Since we have some autocorrelation, we compute the standards error using the clustering method.

```{r}
#| include: true

fixelog_i <- feols(fml = log(inflation_pct) ~ log(broad_money_pct) + urban_pop_pct + enrol_tertiary_gross + net_enrol_primary + net_enrol_secondary + log(gdp_usd) + log(total_reserve_usd) + log(exports_gdp_pct)| country, data = join_clean, se = "cluster")

fixelog_t <- feols(fml = log(inflation_pct) ~ log(broad_money_pct) + urban_pop_pct + enrol_tertiary_gross + net_enrol_primary + net_enrol_secondary + log(gdp_usd) + log(total_reserve_usd) + log(exports_gdp_pct)| year, data = join_clean, se = "cluster")

fixelog_it <- feols(fml = log(inflation_pct) ~ log(broad_money_pct) + urban_pop_pct + enrol_tertiary_gross + net_enrol_primary + net_enrol_secondary + log(gdp_usd) + log(total_reserve_usd) + log(exports_gdp_pct) | country+year, data = join_clean, se = "cluster")

etable(fixelog_i, fixelog_t, fixelog_it)
```

Accounting for only time fixed-effects, we observe strong evidence against the null hypothesis that 'broad money (%)' coefficient is equal to 0, as well as some evidence against the null hypothesis that 'enrol_tertiary_gross' coefficient is equal to 0.

This difference suggest some kind of interaction between location and time fixed effect, as well as a strong bias on broad money predictors and tertiary enrollement introduced by the location. This means that there are strong time-invariant cofounder effects on those 2 predictors. This makes sense as there big broad money variation between wealth_class.

We have some evidence against the null hypothesis that the coefficient of `broad_money_pct`, `net_enrol_secondary` and `total_reserve_usd` are different from 0. 

Firstly, a 1% increase in the amount of financial instruments possessed by a country would lead to a 0.8571% percent increase in inflation rate. Although one might think that the more broad money a country has, the less inflation it would subject to as associating broad money with wealth seems logical.

Secondly, a 1% percent increase in the amount of enrollement in secondary education (ratio based on enrollmenet / not enrolled) would decrease the inflation rate by 1.614%.

Thirdly, a 1% percent increase in the bank liquid reserve would decrease the inflation rate by 0.299%.
